---
title: "CS 422 - HW 10"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
author: Mark Gameng,
        Illinois Institute of Technology
---

## Part 2.1

### (a)
```{r}
countries <- read.csv("countries.csv", row.names = 1)
countries
```

### (a.i)
```{r}
summary(countries)
```

### (a.ii)
```{r}
# boxplots
boxplot(countries[,1], ylab = "GDP", main = "GDP Boxplot")
boxplot(countries[,2], ylab = "HIV", main = "HIV Boxplot")
boxplot(countries[,3], ylab = "Lifeexp", main = "Lifeexp Boxplot")
boxplot(countries[,4], ylab = "Mil", main = "Mil Boxplot")
boxplot(countries[,5], ylab = "Oilcons", main = "Oilcons Boxplot")
boxplot(countries[,6], ylab = "Pop", main = "Pop Boxplot")
boxplot(countries[,7], ylab = "Tel", main = "Tel Boxplot")
boxplot(countries[,8], ylab = "Unempl", main = "Unempl Boxplot")
```
The two outliers associated with Pop boxplot represents China and India because their population is way above other countries.

### (b)
```{r}
countries.pca <- prcomp(countries, center = TRUE, scale = TRUE)
```

### (c.i)
```{r}
summary(countries.pca)
```
No components explain at least 90% of the variance. The highest has variance of 0.531.

### (c.ii)
```{r}
library(factoextra)
fviz_eig(countries.pca)
```
### (c.iii)
Based on the screeplot, I would use 4 or 5 components for modeling if I were to engage in a feature reduction tastk.

### (d)
```{r}
countries.pca$rotation
```

### (d.i)
PC1 is positively correlated with GDP, Lifeexp, Oilcons, and Tel. It is negatively correlated with HIV, Mil, Pop, and Unempl. Based on this, my expectation of PC1 is that it is focused on the money side of things because positively correlated to GDP, Oilcons, Tel and Lifeexp which, in general, the more money, the higher the life expectancy. GDP, Lifeexp, Oilcons, and Tel all are high values, being close to 0.45.

### (d.ii)
PC2 is positively correlated with GDP, Lifeexp, Mil, Oilcons, Pop, Tel. It is negatively correlated with HIV, Unempl. Based on this, my expectations of PC2 is that it is similar to PC1 but more general as it correlates more to Mil and Pop. It is also more negatively correlated to HIV and Unempl (-.193 vs -.328 and -.21 vs -.5). Mil and Pop have the most magnitude and is the most relevant, 0.5671 and 0.5419.


### (e)
```{r}
biplot(countries.pca, choices = 1:2)
```

### (e.i)
```{r}
countries.pca$x[c("Brazil", "UK", "Japan"),1:2]
```

### (e.ii)
Brazil has PC1 = -2.04 and PC2 = -1.04. This makes sense because Brazil has pretty low attributes across the board which resulted in negative values for PC1 and PC2.

Japan has PC1 = 2.01 and PC2 = -0.0574. This makes sense because while Japan has a pretty high GDP, Lifeexp, Oilcons, and Tel, which results in a quite high positive value for PC1, but has a pretty low Mil and Pop which results in a pretty low and negative value for PC2.

UK has PC1 = 1.46 and PC2 = 0.8345. This makes sense because UK has quite high GDP, Lifeexp, Oilcons, and Tel but low Mil and Pop which makes PC2 lower.

## Part 2.2

### User Profiles
```{r}
library(lsa)
users <- read.csv("ratings.csv")
movies <- read.csv("movies.csv")
my_id <- 20419026 %% 671
profile.dfu <- users[users$userId == my_id,]
#profile.dfu
profile.dfm <- movies[movies$movieId %in% profile.dfu$movieId,]
#profile.dfm
profile.tuser <- data.frame()
names <- c("Action", "Adventure", "Animation",
           "Children", "Comedy", "Crime", "Documentary",
           "Drama", "Fantasy", "Film-Noir", "Horror", 
           "IMAX", "Musical", "Mystery", "Romance", 
           "Sci-Fi", "Thriller", "War", "Western", 
           "(no genres listed)")
for (k in names) profile.tuser[[k]] <- as.numeric()
#profile.dfm
for (name in row.names(profile.dfm)){
  genres <- profile.dfm[name, c("genres")]
  genres <- strsplit(genres, "\\|")
  name <- if (name == "Sci-Fi") "Sci.Fi" else name
  for (genre in genres){
    profile.tuser[as.character(profile.dfm[name, c("movieId")]), as.character(genre)] = 1
  }
}
#profile.tuser
profile.user <- colSums(profile.tuser[,], na.rm = TRUE)
profile.user <- profile.user/nrow(profile.dfm)
profile.user
```

### Movie Profiles
```{r}
profile.dfms <- movies[sample(nrow(movies), 10),]
profile.tdfms <- data.frame()
names <- c("Action", "Adventure", "Animation",
           "Children", "Comedy", "Crime", "Documentary",
           "Drama", "Fantasy", "Film-Noir", "Horror", 
           "IMAX", "Musical", "Mystery", "Romance", 
           "Sci-Fi", "Thriller", "War", "Western", 
           "(no genres listed)")
for (k in names) profile.tdfms[[k]] <- as.numeric()
for (name in row.names(profile.dfms)){
  genres <- profile.dfms[name, c("genres")]
  genres <- strsplit(genres, "\\|")
  #cat(unlist(genres), "\n")
  for (genre in genres){
    profile.tdfms[as.character(profile.dfms[name, c("movieId")]), as.character(genre)] = 1
  }
}
#profile.dfms
profile.tdfms
```

### Recommendations
```{r}
profile.tdfms[is.na(profile.tdfms)] <- 0
out.movieId <- c()
out.movieName <- c()
out.similarity <- c()
i <- 1
for (movie in row.names(profile.tdfms)){
  mp <- unlist(profile.tdfms[movie,])
  #typeof(unlist(profile.tdfms[movie,]))
  #cat(unlist(profile.tdfms[movie,]), typeof(unlist(profile.tdfms[movie,])), "\n")
  sim <- lsa::cosine(profile.user, mp)
  out.movieId[i] <- movie
  out.movieName[i] <- movies[movies$movieId == movie, c("title")]
  out.similarity[i] <- sim
  i <- i + 1
}
output <- data.frame(MovieId = out.movieId, MovieName = out.movieName, 
                     Similarity = out.similarity)
output <- output[order(output$Similarity, decreasing = TRUE),]
cat("User ID", my_id, "chose the following 10 movies:", row.names(profile.tdfms), "\n")
cat("Of these, the following 5 movies are reccomended:", "\n")
head(output, 5)
```